---
title: "Appendix"
author:
  - 'Sokpheanal Huynh'
  - 'Mark Newman'
date: "2024/05/15"
editor_options: 
  chunk_output_type: console
---
add commentary
```{r message = FALSE, warning = FALSE, echo = FALSE}
here::i_am('Appendix.rmd')
library(dplyr)
library(tidyr)
library(lubridate)
library(knitr)
library(kableExtra)
library(ggplot2)
library(nortest)
library(here)

rm(list = ls())
options(scipen = 999)
knitr::opts_chunk$set(results = 'hold', echo = FALSE)
```

* [t1](manual extract)
* [t3](https://github.com/sokpheanal/FinnHub)

```{r message = FALSE, warning = FALSE, echo = FALSE}
leg <- read.csv(here('data/legislators.csv'))
trans <- read.csv(here('data/all_available_transactions.csv'))
```

```{r}
str(leg)
str(trans)
View(leg)
View(trans)
```

```{r message = FALSE, warning = FALSE, echo = FALSE}
leg <-
  leg %>%
  select(-X)
leg <-
  leg %>%
  mutate(
    position = factor(position),
    gender = factor(gender, levels = c('M', 'F'), labels = c('Male', 'Female')),
    state = factor(state),
    party = factor(party))

str(leg)
```

```{r}
trans <-
  trans %>%
  select(-assetName, -ownerType)
str(trans)

trans <-
  trans %>%
  mutate(
    filingDate = mdy(filingDate),
    transactionDate = mdy(transactionDate))
str(trans)
View(trans)
```


**Question (Exploration)**: How many members of congress report trades?

```{r}
length(unique(trans$name))
```

suspect, but lets move forward

**Question (Exploration)**: What are the names of members of congress that made trades, but do not appear in our crosswalk?

```{r}
t1 <-
  trans %>%
  select(name, position) %>%
  unique() %>%
  left_join(leg, by = c('name' = 'name', 'position' = 'position'))

nrow(t1[is.na(t1$party),]) == 0
```

Make final dataset b/c nrow == 0

```{r}
data <-
  trans %>%
  left_join(leg, by = c('name' = 'name', 'position' = 'position'))
nrow(data) == nrow(trans)
nrow(data[is.na(data$party),]) == 0
View(data)

rm(list = ls()[!(ls() %in% 'data')])
```

**Question (Exploration)**: How many unique tickers?

```{r}
length(unique(data$symbol))
```

**Question (Exploration)**: Basic description of the data

```{r}
data <-
  data %>%
  mutate(lag = (filingDate - transactionDate)/ddays(x = 1))
str(data)
View(data)
```


```{r}
t1 <- 
  data %>%
  select(gender)

t1 %>%
  group_by(gender) %>%
  summarise(
    count = n()) %>%
  arrange(desc(count)) %>%
  kable(
    caption = 'Trades Count by Gender',
    col.names = c('Gender', 'Count')) %>%
  kable_styling()

ggplot(t1,
       aes(x = gender, fill = gender)) +
  theme_bw() +
  geom_bar() +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = 'Figure 1. Trades Count by Gender')
```

```{r}
t2 <- 
  data %>%
  select(party)

t2 %>%
  group_by(party) %>%
  summarise(
    count = n()) %>%
  arrange(desc(count)) %>%
  kable(
    caption = 'Trades Count by Political Party',
    col.names = c('Party', 'Count')) %>%
  kable_styling()

ggplot(t2, aes(x = party, fill = party)) +
  theme_bw() +
  geom_bar() + 
  scale_fill_brewer(palette = "Dark2") +
  labs(title = 'Figure 2. Trades Count by Party')
```

```{r}
t3 <- 
  data %>%
  select(state)

t3 %>%
  group_by(state) %>%
  summarise(
    count = n()) %>%
  arrange(desc(count)) %>%
  kable(
    col.names = c('State', 'Count')) %>%
  kable_styling()
```

```{r}
t4 <- 
  data %>%
  select(lag) %>%
  filter(!is.na(lag))

t4 %>%
  summarise(
    mean = mean(lag),
    median = median(lag),
    sd = sd(lag),
    iqr = IQR(lag)) %>%
  kable(
    col.names = c('Mean', 'Median', 'SD', "IQR"),
    digits = 2) %>%
  kable_styling() %>%
  add_header_above(header = c('Lag' = 4))

ggplot(t4, aes(y = lag)) +
  theme_bw() +
  geom_boxplot() +
  labs(title = 'Overall Lag Descriptive Statistics',
       caption = 'Figure 3. Overall Lag Descriptive Statistics')
```

**Questions**

Who files fastest: (tests of `lag`) 

* M vs F (`gender`)
* D vs R (`party`)
* what state is the most responsive (`state`)
   
```{r}
t1 <- 
  data %>%
  select(lag, gender) %>%
  filter(!is.na(lag))

t1 %>%
  group_by(gender) %>%
  summarise(
    mean = mean(lag),
    median = median(lag),
    sd = sd(lag),
    iqr = IQR(lag)) %>%
  arrange(desc(median)) %>%
  kable(
    caption = 'Lag by Gender',
    col.names = c('Gender', 'Mean', 'Median', 'SD', "IQR"),
    digits = 2) %>%
  kable_styling() %>%
  add_header_above(header = c(' ' = 1, 'Lag' = 4))

ggplot(t1, aes(x = lag, color = gender)) +
  theme_bw() +
  geom_density() +
  scale_color_brewer(palette = "Dark2") +
  labs(title = 'Lag by Gender')

lillie.test(t1$lag[t1$gender == 'Male'])
lillie.test(t1$lag[t1$gender == 'Female'])

wilcox.test(lag ~ gender, data = t1)
```

`lillie.test()` demonstrates non-normality, choose non-parametric test.
`Wilcox.test()`: p is less than 0.05 so we reject the null hypothesis and say that there is evidence to suggest a difference in group median.


```{r}
t2 <- 
  data %>%
  select(lag, party) %>%
  filter(!is.na(lag))

t2 %>%
  group_by(party) %>%
  summarise(
    mean = mean(lag),
    median = median(lag),
    sd = sd(lag),
    iqr = IQR(lag)) %>%
  arrange(desc(median)) %>%
  kable(
    caption = 'Lag by Party',
    col.names = c('Party', 'Mean', 'Median', 'SD', 'IQR'),
    digits = 2) %>%
  kable_styling() %>%
  add_header_above(header = c(' ' = 1, 'Lag' = 4))

ggplot(t2, aes(x = lag, color = party)) +
  theme_bw() +
  geom_density() +
  scale_color_brewer(palette = "Dark2")

lillie.test(t2$lag[t2$party == 'Democratic'])
lillie.test(t2$lag[t2$party == 'Republican'])

wilcox.test(lag ~ party, data = t2)
```

`lillie.test()` demonstrates non-normality, choose non-parametric test.
`Wilcox.test()`: p is less than 0.05 so we reject the null hypothesis and say that there is evidence to suggest a difference in group median.

```{r}
t3 <- 
  data %>%
  select(lag, state) %>%
  filter(!is.na(lag))

t3 %>%
  group_by(state) %>%
  summarise(
    mean = mean(lag),
    median = median(lag),
    sd = sd(lag),
    iqr = IQR(lag)) %>%
  arrange(desc(median)) %>%
  kable(
    caption = 'Lag by State',
    col.names = c('State', 'Mean', 'Median', 'SD', 'IQR'),
    digits = 2) %>%
  kable_styling() %>%
  add_header_above(header = c(' ' = 1, 'Lag' = 4))
```
